\pdftitle

### 2017-05-18 maxNt.tex
discussion

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-18T14:02:33+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

Suppose we want to constrain the sample expectations of a vector function
$\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\ at the population level,
denoted by $\yHa$, gives
\begin{equation}
  \label{eq:app_maxent_pop}
  \p(\yXf = \yRf \| \yHa)  = \yK\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
and then by marginalization with \eqn~\eqref{eq:marginal}
\begin{equation}
  \label{eq:app_maxent_pop_marg}
  \p(\yxs = \yrs \| \yHa)  = \yK\, 
  \sum_{N\yRf=0}^N \ypp(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
where $\yK$ is a normalization constant and
$\yL\T=(\varLambda_1,\dotsc,\varLambda_m)\T$ are Lagrange multipliers such
that
\begin{equation}
  \label{eq:app_maxent_pop_constraints}
\yc = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N \yf(\yrs) \ypp(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}
Application of \me\ at the sample level, denoted by $\yHb$, gives
\begin{equation}
  \label{eq:app_maxent_sample}
  \p(\yxs=\yrs \|\yHb)
  =\yk\,
  \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)]
\end{equation}
where $\yk$ is a normalization constant and $\yl\T$ are Lagrange
multipliers such that
\begin{equation}
  \label{eq:app_maxent_sample_constraints}
\yc = \yk\sum_{n\yrs=0}^n  \yf(\yrs) \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)].
\end{equation}

@@

Yet they cannot be exactly equal, because their equality would require the
Lagrange multipliers $\yL$ and $\yl$ to satisfy equations
\eqref{eq:app_maxent_pop_constraints},
\eqref{eq:app_maxent_sample_constraints}, and
$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ -- that is, $2m+n$
equations in $m$ unknowns. This could be possible, if at all, only for very
special choices of constraints functions $\yf$ and values $\yc$.

@@

A perfunctory intuitive reasoning seems insufficient for clarifying this
point. Let's express it via the probability calculus. Suppose we do not
know whether the sample is really part of a larger population; we do not
know if $N=n$ or how large is $N$ otherwise. Call this state of ignorance
$\yHd$. From the point of view of the probability calculus, this ignorance
about $N$ is expressed by assigning a probability distribution
$\p(N \| \yHd)$ that vanishes if $N<n$, since we know that $N\ge n$.
Maintaining our assumption of symmetric ignorance, probability assignments
that do not assume a specific value of $N$ are then obtained by
multiplication of all $N$-dependent probabilities by $\p(N \| \yHd)$ and
subsequent marginalization over $N$ -- technically speaking, $N$ is a
\emph{nuisance parameter} that we eliminate. The probability obtained from
\me\ at the population level, \eqn~\eqref{eq:app_maxent_pop_marg} then
generalizes to
\begin{equation}
  \label{eq:app_maxent_pop_marg_unknown_N}
  \p(\yxs = \yrs \| \yHd)  = \sum_N \biggl\{\yK_N\, 
  \sum_{N\yRf=0}^N \ypp_N(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[{\yL_N}\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
  \,\p(N \| \yHd).
\end{equation}
This is a formidable expression. But the answer to our question, whether
the usual \me\ at the sample level~\eqref{eq:app_maxent_sample} does not
assume anything about a larger population, translates into the precise
mathematical question: are the
distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
\eqref{eq:app_maxent_sample} equal, for some choice of $\p(N \| \yHd)$? We
leave this mathematical problem for future work.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2017-05-18T14:18:19+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

Suppose we want to constrain the sample expectations of a vector function
$\yf = (f_1,\dotsc,f_m)$ to the vector values $\yc=(c_1,\dotsc,c_m)$, that
is, $\expeb{\yf(\yxs)}=\yc$. Application of \me\
\cite{sivia1996_r2006,meadetal1984} at the population level, denoted by
$\yHa$, gives
\begin{equation}
  \label{eq:app_maxent_pop}
  \p(\yXf = \yRf \| \yHa)  = \yK\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
and then by marginalization with \eqn~\eqref{eq:marginal}
\begin{equation}
  \label{eq:app_maxent_pop_marg}
  \p(\yxs = \yrs \| \yHa)  = \yK\, 
  \sum_{N\yRf=0}^N \ypp(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr],
\end{equation}
where $\yK$ is a normalization constant and
$\yL\T=(\varLambda_1,\dotsc,\varLambda_m)\T$ are Lagrange multipliers such
that
\begin{equation}
  \label{eq:app_maxent_pop_constraints}
\yc = \yK\sum_{n\yrs=0}^n \sum_{N\yRf=0}^N \yf(\yrs) \ypp(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[\yL\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp(\yrs\|\yRf)\Bigr].
\end{equation}
Application of \me\ at the sample level, denoted by $\yHb$, gives
\begin{equation}
  \label{eq:app_maxent_sample}
  \p(\yxs=\yrs \|\yHb)
  =\yk\,
  \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)]
\end{equation}
where $\yk$ is a normalization constant and $\yl\T$ are Lagrange
multipliers such that
\begin{equation}
  \label{eq:app_maxent_sample_constraints}
\yc = \yk\sum_{n\yrs=0}^n  \yf(\yrs) \binom{n}{n\yrs}
  \exp[\yl\T \yf(\yrs)].
\end{equation}

@@

Yet they cannot be exactly equal, because their equality would require the
Lagrange multipliers $\yL$ and $\yl$ to satisfy equations
\eqref{eq:app_maxent_pop_constraints},
\eqref{eq:app_maxent_sample_constraints}, and
$\p(\yxs = \yrs \| \yHa) = \p(\yxs = \yrs \| \yHb)$ -- that is, $2m+n$
equations (normalization is taken care of) in $2m$ unknowns. A solution can
exist, if at all, only for very special choices of constraints functions
$\yf$ and values $\yc$.

@@

A perfunctory intuitive reasoning seems insufficient for clarifying this
point. Let's express it in the language of the probability calculus.
Suppose we do not know whether the sample is really part of a larger
population: we do not know whether $N=n$ or how large is $N$ otherwise.
Call this state of ignorance $\yHd$. From the point of view of the
probability calculus, this ignorance about $N$ is expressed by assigning a
probability distribution $\p(N \| \yHd)$ that vanishes if $N<n$, since we
know that $N\ge n$. Maintaining our assumption of symmetric ignorance,
probability assignments that do not assume a specific value of $N$ are then
obtained by multiplication of all $N$-dependent probabilities by
$\p(N \| \yHd)$ and subsequent marginalization over $N$ -- technically
speaking, $N$ is a \emph{nuisance parameter} that we eliminate
\cite{jaynes1994_r2003,lindley1965b_r2008,bernardoetal1994_r2000}. The
probability obtained from \me\ at the population level,
\eqn~\eqref{eq:app_maxent_pop_marg} then generalizes to
\begin{equation}
  \label{eq:app_maxent_pop_marg_unknown_N}
  \p(\yxs = \yrs \| \yHd)  = \sum_N \biggl\{\yK_N\, 
  \sum_{N\yRf=0}^N \ypp_N(\yrs \|\yRf)
\,
  \binom{N}{N \yRf}\exp\Bigl[{\yL_N}\T
  \sum_{n\yrs=0}^n \yf(\yrs)\ypp_N(\yrs\|\yRf)\Bigr]\biggr\}
  \,\p(N \| \yHd).
\end{equation}
This is a formidable expression. But the answer to our question, whether
the usual \me\ at the sample level~\eqref{eq:app_maxent_sample} does not
assume anything about a larger population, translates into the precise
mathematical question: are the
distributions~\eqref{eq:app_maxent_pop_marg_unknown_N} and
\eqref{eq:app_maxent_sample} equal, for some choice of $\p(N \| \yHd)$? We
leave this mathematical problem for future work.
